---
title: "gtsummary"
output:
  html_document:
    df_print: paged
---
```{r setup, warning = FALSE, message = FALSE}
library(tidyverse)
library(gtsummary)
library(palmerpenguins)
```

**Example 1**

```{r data import, message=FALSE, warning=FALSE}
draco_ecomorph1 <- read_csv("draco_ecomorph.csv")
```

```{r example 1 make the table}
#this function works best with datasets without NA values

table_Draco1 <- draco_ecomorph1 %>%
  select(species, sex, snout_vent_length_mm, perch_height_cm) %>% #retain variables used in summaries
  filter(species %in% c("bimaculatus", "cornutus", "melanopogon")) %>%
  tbl_summary(by = species,
              missing = "no", # remove missing values (NAs). "ifany" will include a missing values row in table.
              digits = all_continuous() ~ 1, #number of displayed digits for continuous variables
              label = list(snout_vent_length_mm ~ "SVL (mm)", #change column labels in table
                           perch_height_cm ~ "Perch Height (cm)",
                           sex ~ "Sex"),
              statistic = list(all_continuous() ~ "{mean} ({sd})", #can use any character for parentheses
                               all_categorical() ~ "{n}")) %>% #count obs. for categorical variables
  #add_p(snout_vent_length_mm ~ "aov") %>% CANT FIGURE OUT HOW TO FORCE ANOVA
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #number of digits displayed for p-values
  modify_caption("Table 1. Lizard and Habitat Characteristics") %>%
  #modify_header(update = all_stat_cols() ~ "**{level}**") %>% #headers to only contain species name
  modify_header(update = list(
  label ~ '',
  stat_1 ~ '**D. bimaculatus**',
  stat_2 ~ '**D. cornutus**',
  stat_3 ~ '**D. melanopogon**',
  p.value ~ '**P-value**'
))

#show_header_names(table_Draco) 
#lets you preview code for changing header names and formatting  
                
table_Draco1
```

```{r save Draco table}
#table_Draco1 %>% 
  #as_gt() %>%   # convert to gt table for additional formatting options and saving
  #gt::gtsave(filename = "draco_ecomorph_table.pdf") #save file
```
If data have missing values that make groups unbalanced, might have to make more than one table and combine them.
Parts of a gt table: https://gt.rstudio.com/articles/intro-creating-gt-tables.html

Advanced Customization

It's natural a {gtsummary} package user would want to customize the aesthetics of the table with one or more of the many {gt} functions available.

Every function in {gt} is available to use with a {gtsummary} object.

1. Create a {gtsummary} table.
2. Convert the table to a {gt} object with the as_gt() function.
3. Continue formatting as a {gt} table with any {gt} function.

**Example 2**

```{r palmer penguins data}
table_penguins <- penguins %>%
  select(species, sex, bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g) %>% #retain variables used in summaries
  tbl_summary(by = species,
              missing = "no", # remove missing values (NAs). "ifany" will include a missing values row in table.
              digits = all_continuous() ~ 1, #number of displayed digits for continuous variables
              label = list(bill_length_mm ~ "Bill Length (mm)", #change column labels in table
                           bill_depth_mm ~ "Bill Depth (mm)",
                           flipper_length_mm ~ "Flipper Length (mm)",
                           body_mass_g ~ "Body Mass (g)",
                           sex ~ "Sex"),
              statistic = list(all_continuous() ~ "{mean} ({sd})", #can use any character for parentheses
                               all_categorical() ~ "{n}")) %>% #count obs. for categorical variables
  add_p(pvalue_fun = ~style_pvalue(.x, digits = 2)) %>% #number of digits displayed for p-values
  modify_caption("Table 2. Palmer Penguin Morphometrics") %>%
  modify_header(update = all_stat_cols() ~ "**{level}**") %>% #headers to only contain species name
  as_gt() %>% #have to convert to gt object for some types of formatting
  gt::tab_options(heading.align = "left") #can't get to work; most journals want table caption separate anyway
#table caption appears left aligned after it is knit to html.

table_penguins
```
If more formatting is required, likely have to convert `gtsummary` object to a `gt` object and modify with `gt` package.

Other worked examples:
https://cran.r-project.org/web/packages/gtsummary/vignettes/gallery.html


